Draft Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-03-22 draftfinalreport_v2.tex
Hopefully final draft...

This large number of categories, some with small population sizes, leads to both overdispersion and zeroinflation, as there are many categories with zero cases in, particularly in the adult age classes. Furthermore, initial preliminary analyses, including multi-- and univariate analyses (not shown) suggested little effect of \textit{individual} NZDep classifications and several higher order interactions, and therefore we reduced the number of NZDep categories from ten to two: NZDep 1--5 (least deprived) and NZDep 6--10 (most deprived) (\href{http://www.otago.ac.nz/wellington/research/hirp/otago020194.html}{www.otago.ac.nz}). We also incorporated the 65+ age classes into the 25--64 age category, to make a 25+ age category. By doing so, we reduce the zeroinflation present in the data.

The Prioritized Ethnicities for cases are:  European; Maori; Pacific Peoples; Asian; Middle Eastern/Latin American/African (MLA); Other Ethnicity; Residual Categories. For the analyses in this report only the first five are used, as these categories cover the overwhelming number of cases, with only 1.9\% (22/1137) of cases having no Prioritized Ethnicity (see \emph{None}, \autoref{table:none}).

% \begin{table}[hbtp]\scriptsize
% <<results=tex,echo=FALSE>>=# This chunk uses "results=tex"
% latex(ttyr, file="",table.env=FALSE,rowname=NULL)
% @
% \caption{Absolute number of measles cases with no ethnicity recorded for specific age and socio-economic deprivation (NZDep) categories from 2007-2014.}
% \label{table:none}
% \end{table}

\begin{sidewaysfigure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{percap_ordert.pdf}
\end{center}
\caption{test}
\label{fig:percap_all}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{case_ordert.pdf}
\end{center}
\caption{test}
\label{fig:cases_all}
\end{sidewaysfigure}

Two way interactions are shown in \autoref{sec:app} (\autoref{fig:CaseAgeNzdep}, \autoref{fig:CaseNzdepEth}, \autoref{fig:CaseAgeEth}, \autoref{fig:PerCapAgeNzdep}, \autoref{fig:PerCapNzdepEth}, \autoref{fig:PerCapAgeEth}).

% \begin{table}[hbtp]\scriptsize
% \caption{Per capita measles rates}
% <<results=tex,echo=FALSE>>=
% latex(PerCap, file="", table.env=FALSE,rowname=NULL)
% @
% \label{table:percap_rates}
% \end{table}

\begin{table}[hbtp]\footnotesize
\caption{Proportions of the na\"{\i}ve population the would be covered if vaccinated by age groups}
%latex.default(propvCohort, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{ll}
\hline\hline
\multicolumn{1}{c}{Age range}&\multicolumn{1}{c}{Proportion}\tabularnewline
\hline
2-5&0.04\tabularnewline
5-11&0.15\tabularnewline
5-16&0.27\tabularnewline
5-18&0.31\tabularnewline
11-16&0.14\tabularnewline
11-18&0.18\tabularnewline
0-29&0.72\tabularnewline
0-44&0.89\tabularnewline
\textgreater8&0.8\tabularnewline
\hline
\end{tabular}\end{center}\label{table:vaccine_cohort}
\end{table}

The numbers of cases per category and population sizes for the complete data set from 2007-2014 can be seen in \autoref{table:percap}. Subsequent regression analyses (not shown) also suggested that the MLA category was over-- or underrepresented in per capita rates given the very small sample sizes for this classification (\autoref{fig:percap}, \autoref{table:percap}), leading to very large standard error in regression analyses. However, there are numerous issues with the data for MLA ethnicity category above and beyond the small population sizes (see \autoref{table:percap}), thus creating issues for estimating the denominator data for this group (University of Otago, personal communication). Thus, we removed this grouping for our subsequent analyses and are left with Asian, European, Maori and Pacific as Prioritized Ethnicities. This left us with 1102/1115 (99\%) of the measles cases with Prioritized Ethnicity recorded from 2007, and 1102/1137 (97\%) of all measles cases recorded since 2007 (\autoref{table:percap}).

@@

\caption{Measles cases (Cases), population sizes (Population) and per capita rates per 10,000 (Per capita) of measles in specific age (Age), ethnicity (Ethnicity) and socio-economic deprivation (NZDep) categories from 2007-2014}

@@

\end{tabular}\end{center}\label{table:percap}

@@

@@

The distribution of the measles cases per category used in the regression analyses are in \autoref{fig:histcasecat}.
\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{Cases_regmodel.pdf}
\end{center}
\caption{Distribution of measles cases per category (\autoref{table:percap}) used in the final regression model (\autoref{eq:reg})}
\label{fig:histcasecat}
\end{figure}

The predicted values from the regression model plotted against the reported cases are shown in \autoref{fig:predict}, and the residuals are shown in \autoref{fig:resid}.
\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{Cases_regmodel_prediction.pdf}
\end{center}
\caption{Regression model (\autoref{eq:reg}) predictions plotted against the case data (\autoref{table:percap})}
\label{fig:predict}
\end{figure}

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{Cases_regmodel_resid.pdf}
\end{center}
\caption{Histogram of residuals from the regression model (\autoref{eq:reg})}
\label{fig:resid}
\end{figure}

The significance of the different predictor variables can be seen in the ANOVA results (\autoref{table:anova}).

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Sun Mar 22 01:29:41 2015
\begin{tabular}{lrrrrrr}
  \hline
 & Df & Deviance & Resid. Df & Resid. Dev & F & Pr($>$F) \\ 
  \hline
NULL &  &  & 59 & 1532.80 &  &  \\ 
  Age & 4 & 1345.17 & 55 & 187.63 & 204.01 & 0.0000 \\ 
  Ethnicity & 3 & 19.88 & 52 & 167.75 & 4.02 & 0.0141 \\ 
  NZDep & 2 & 17.10 & 50 & 150.65 & 5.19 & 0.0102 \\ 
  Age:Ethnicity & 12 & 80.80 & 38 & 69.85 & 4.09 & 0.0004 \\ 
   \hline
\end{tabular}\caption{Significance of different predictor variables for measles risk factors from 2007-2014, estimated using \autoref{eq:reg} and a quasipoisson error structure}
\label{table:anova}
\end{table}

A summary of the regression model (\autoref{eq:reg}) with the individual effects and the statistical support for the estimated coefficients can be seen in \autoref{table:regression_riskfactors}.

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Sun Mar 22 01:29:41 2015
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -6.1644 & 0.1267 & -48.65 & 0.0000 \\ 
  Age2-4 & -1.0351 & 0.1972 & -5.25 & 0.0000 \\ 
  Age5-17 & -0.9954 & 0.1368 & -7.28 & 0.0000 \\ 
  Age18-24 & -1.7204 & 0.1926 & -8.93 & 0.0000 \\ 
  Age25+ & -3.1697 & 0.1622 & -19.55 & 0.0000 \\ 
  EthnicityAsian & 0.2464 & 0.3168 & 0.78 & 0.4415 \\ 
  EthnicityMaori & 0.2081 & 0.2121 & 0.98 & 0.3327 \\ 
  EthnicityPacific & 1.2200 & 0.2134 & 5.72 & 0.0000 \\ 
  NZDep4-7 & -0.2609 & 0.0948 & -2.75 & 0.0090 \\ 
  NZDep8-10 & -0.3043 & 0.1040 & -2.93 & 0.0058 \\ 
  Age2-4:EthnicityAsian & -2.3148 & 1.3319 & -1.74 & 0.0903 \\ 
  Age5-17:EthnicityAsian & -1.2453 & 0.4566 & -2.73 & 0.0096 \\ 
  Age18-24:EthnicityAsian & -1.0187 & 0.5743 & -1.77 & 0.0841 \\ 
  Age25+:EthnicityAsian & 0.2344 & 0.4298 & 0.55 & 0.5887 \\ 
  Age2-4:EthnicityMaori & -0.1013 & 0.3644 & -0.28 & 0.7826 \\ 
  Age5-17:EthnicityMaori & -0.1032 & 0.2521 & -0.41 & 0.6846 \\ 
  Age18-24:EthnicityMaori & -0.5723 & 0.4409 & -1.30 & 0.2022 \\ 
  Age25+:EthnicityMaori & -1.0349 & 0.5127 & -2.02 & 0.0506 \\ 
  Age2-4:EthnicityPacific & -0.9798 & 0.4676 & -2.10 & 0.0429 \\ 
  Age5-17:EthnicityPacific & -1.5709 & 0.3312 & -4.74 & 0.0000 \\ 
  Age18-24:EthnicityPacific & -1.0354 & 0.5003 & -2.07 & 0.0453 \\ 
  Age25+:EthnicityPacific & -0.6629 & 0.4279 & -1.55 & 0.1296 \\ 
   \hline
\end{tabular}\caption{Summary of the regression model results for measles risk factors for cases from 2007-2014}
\label{table:regression_riskfactors}
\end{table}
\vspace{5mm} %5mm vertical space

Apart from over-representation of some MLA categories discussed above and not included here, the results of the regression model suggest that age is a strong predictor of being a measles case. Indeed, all age categories are significantly less likely to be measles cases compared to 0-2 year olds, and the likelihood generally decreases with age (see also \autoref{fig:ageinyears}).

The majority of measles cases are among people of European ancestry (\autoref{table:percap} and \autoref{fig:ageinyears}), however, people of Pacific origin are also over-represented as measles cases on a per capita basis ($\beta$ = 1.16, standard error (SE) = 0.2, p--value < 0.0001). Those people in NZDep levels 6-10 are under-represented ($\beta$ = -0.21, SE = 0.085, p--value = 0.02), and there are some other age:ethnicity classes that are significantly less represented in the data compared to Europeans in those ages classes, particularly in the 6--17 age classes. In later measles outbreaks (since 2007) there has been a shift in the distribution of ages infected. The very young (<2 years old) are still most likely to be infected, but school aged children and older teenagers are more likely to be represented then the under tens (\autoref{fig:ageinyears} and \autoref{fig:ageandvac}, \autoref{table:sero}, and \autoref{sub:popim}). This pattern suggests that improving vaccination coverage in the young is reducing the burden of measles in those age categories (\autoref{table:percap}). Interestingly the regression results suggest risk of measles in the 6--17 year age category was greater for Europeans and Maori.

@@

\begin{table}
\caption{Highest measles incidence per million (2012)}
%latex.default(incidence, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{incidence}\tabularnewline
\hline
Equatorial Guinea&$1617$\tabularnewline
Nauru&$1100$\tabularnewline
Democratic Republic of the Congo&$1096$\tabularnewline
Somalia&$ 979$\tabularnewline
Djibouti&$ 824$\tabularnewline
Sudan&$ 786$\tabularnewline
Burkina Faso&$ 447$\tabularnewline
Romania&$ 342$\tabularnewline
Ukraine&$ 280$\tabularnewline
Sudan&$ 229$\tabularnewline
Angola&$ 214$\tabularnewline
Monaco&$ 132$\tabularnewline
Nepal&$ 122$\tabularnewline
Sierra Leone&$ 113$\tabularnewline
Afghanistan&$  93$\tabularnewline
Yemen&$  91$\tabularnewline
Lesotho&$  87$\tabularnewline
Qatar&$  78$\tabularnewline
Thailand&$  78$\tabularnewline
Malaysia&$  64$\tabularnewline
Zambia&$  64$\tabularnewline
Ghana&$  64$\tabularnewline
Indonesia&$  63$\tabularnewline
Congo&$  60$\tabularnewline
Uganda&$  56$\tabularnewline
Libyan Arab Jamahiriya&$  52$\tabularnewline
Ethiopia&$  47$\tabularnewline
Pakistan&$  45$\tabularnewline
Myanmar&$  41$\tabularnewline
Nigeria&$  38$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:incidence12}
\end{table}

\begin{table}
\caption{Lowest national measles vaccine cover (\%), 2012}
%latex.default(vaccinecover, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{cover}\tabularnewline
\hline
Equatorial Guinea&$34$\tabularnewline
Somalia&$49$\tabularnewline
Lesotho&$60$\tabularnewline
Central African Republic&$65$\tabularnewline
Papua New Guinea&$67$\tabularnewline
Chad&$69$\tabularnewline
Haiti&$69$\tabularnewline
South Sudan&$70$\tabularnewline
Gabon&$71$\tabularnewline
Yemen&$71$\tabularnewline
Benin&$72$\tabularnewline
Lao People's Democratic Republic&$72$\tabularnewline
Togo&$72$\tabularnewline
Suriname&$73$\tabularnewline
Timor-Leste&$73$\tabularnewline
Paraguay&$74$\tabularnewline
Mauritania&$75$\tabularnewline
Namibia&$76$\tabularnewline
Eritrea&$77$\tabularnewline
Marshall Islands&$78$\tabularnewline
Nigeria&$78$\tabularnewline
Syrian Arab Republic&$78$\tabularnewline
Ukraine&$79$\tabularnewline
Congo&$80$\tabularnewline
Ethiopia&$80$\tabularnewline
Liberia&$80$\tabularnewline
Afghanistan&$81$\tabularnewline
Cameroon&$82$\tabularnewline
Mozambique&$82$\tabularnewline
Senegal&$82$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:cover12}
\end{table}

\begin{table}
\caption{Non-New Zealander travel and immigration numbers (2012)}
%latex.default(im, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{immigration}\tabularnewline
\hline
Australia&$809775$\tabularnewline
United Kingdom&$306177$\tabularnewline
China&$256036$\tabularnewline
United States&$194438$\tabularnewline
Japan&$ 86676$\tabularnewline
Germany&$ 83608$\tabularnewline
Korea, Republic of&$ 73459$\tabularnewline
France&$ 71448$\tabularnewline
India&$ 69038$\tabularnewline
Canada&$ 54981$\tabularnewline
Malaysia&$ 50506$\tabularnewline
Fiji&$ 46723$\tabularnewline
Netherlands&$ 33431$\tabularnewline
South Africa&$ 30600$\tabularnewline
Tonga&$ 26717$\tabularnewline
Singapore&$ 25460$\tabularnewline
Ireland&$ 25036$\tabularnewline
Samoa&$ 24547$\tabularnewline
Philippines&$ 24527$\tabularnewline
Taiwan&$ 23815$\tabularnewline
Thailand&$ 20258$\tabularnewline
Hong Kong&$ 19103$\tabularnewline
Indonesia&$ 15692$\tabularnewline
Switzerland&$ 14851$\tabularnewline
Brazil&$ 14778$\tabularnewline
Sweden&$ 13719$\tabularnewline
Italy&$ 13388$\tabularnewline
Spain&$ 10104$\tabularnewline
Denmark&$ 10056$\tabularnewline
Russia&$  8103$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:imtravel12}
\end{table}

>>>>

This large number of categories, some with small population sizes, leads to both overdispersion and zeroinflation, as there are many categories with zero cases in, particularly in the adult age classes. Furthermore, initial preliminary analyses, including multi-- and univariate analyses (not shown) suggested little effect of \textit{individual} NZDep classifications and several higher order interactions, and therefore we reduced the number of NZDep categories from ten to three: NZDep 1--3 (least deprived), NZDep 4--7, and NZDep 8--10 (most deprived) (\href{http://www.otago.ac.nz/wellington/research/hirp/otago020194.html}{www.otago.ac.nz}) following discussion with the Ministry regarding the most appropriate categories. We also incorporated the 65+ age classes into the 25--64 age category, to make a 25+ age category. By doing so, we reduce the zeroinflation present in the data.

The Prioritized Ethnicities for cases are:  European; Maori; Pacific Peoples; Asian; Middle Eastern/Latin American/African (MLA); Other Ethnicity; Residual Categories. For the analyses in this report only the first five are used, as these categories cover the overwhelming number of cases, with only 1.9\% (22/1137) of cases having no Prioritized Ethnicity.

\begin{sidewaysfigure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{case_ordert.pdf}
\end{center}
\caption{Measles cases from 2007--2014 (\autoref{table:percap}). The x-axis shows NZDep:Age Range:Prioritized ethnicity. Per capita values are in \autoref{fig:percap_all}}
\label{fig:cases_all}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{percap_ordert.pdf}
\end{center}
\caption{Per capita measles values from 2007--2014 (\autoref{table:percap}). The x-axis shows NZDep:Age Range:Prioritized ethnicity. Absolute cases are in \autoref{fig:cases_all}}
\label{fig:percap_all}
\end{sidewaysfigure}

The numbers of cases per category and population sizes for the complete data set from 2007-2014 can be seen in \autoref{table:percap}. Subsequent regression analyses (not shown) also suggested that the MLA category was over-- or underrepresented in per capita rates given the very small sample sizes for this classification (\autoref{fig:percap_all}, \autoref{table:percap}), leading to very large standard error in regression analyses. However, there are numerous issues with the data for MLA ethnicity category above and beyond the small population sizes (see \autoref{table:percap}), creating issues for estimating the denominator data for this group (University of Otago, personal communication). Thus, we removed this grouping for our subsequent analyses and are left with Asian, European, Maori and Pacific as Prioritized Ethnicities. This left us with 1102/1115 (99\%) of the measles cases with Prioritized Ethnicity recorded from 2007, and 1102/1137 (97\%) of all measles cases recorded since 2007 (\autoref{table:percap}). The breakdown of cases and per capita measles values can be see in \autoref{fig:cases_all} and \autoref{fig:percap_all}. Two way interactions are shown in the Appendix (\autoref{sec:app}, \autoref{fig:CaseAgeNzdep}, \autoref{fig:CaseNzdepEth}, \autoref{fig:CaseAgeEth}, \autoref{fig:PerCapAgeNzdep}, \autoref{fig:PerCapNzdepEth}, \autoref{fig:PerCapAgeEth}).

@@

@@

\end{tabular}\end{center}\caption{Measles cases (Cases), population sizes (Population) and per capita rates per 10,000 (Per capita) of measles in specific age (Age), ethnicity (Ethnicity) and socio-economic deprivation (NZDep) categories from 2007-2014}
\label{table:percap}

@@

\begin{table}[hbtp]\footnotesize
%latex.default(propvCohort, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{ll}
\hline\hline
\multicolumn{1}{c}{Age range}&\multicolumn{1}{c}{Proportion}\tabularnewline
\hline
2-5&0.04\tabularnewline
5-11&0.15\tabularnewline
5-16&0.27\tabularnewline
5-18&0.31\tabularnewline
11-16&0.14\tabularnewline
11-18&0.18\tabularnewline
0-29&0.72\tabularnewline
0-44&0.89\tabularnewline
\textgreater8&0.8\tabularnewline
\hline
\end{tabular}\end{center}\caption{Proportions of the na\"{\i}ve population the would be covered if vaccinated by age groups}
\label{table:vaccine_cohort}
\end{table}

@@

The cases and per capita measles values are in \autoref{fig:cases_all} and \autoref{fig:percap_all}. The distribution of the measles cases per category used in the regression analyses are in \autoref{fig:histcasecat}. The predicted values from the regression model plotted against the reported cases are shown in \autoref{fig:predict}, and the residuals are shown in \autoref{fig:resid}. The significance of the different predictor variables can be seen in the ANOVA results (\autoref{table:anova}). A summary of the regression model (\autoref{eq:reg}) with the individual effects and the statistical support for the estimated coefficients can be seen in \autoref{table:regression_riskfactors}.

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{Cases_regmodel.pdf}
\end{center}
\caption{Distribution of measles cases per category (\autoref{table:percap}) used in the final regression model (\autoref{eq:reg})}
\label{fig:histcasecat}
\end{figure}

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{Cases_regmodel_prediction.pdf}
\end{center}
\caption{Regression model predictions plotted against the case data (\autoref{table:percap}), estimated using \autoref{eq:reg} and a quasipoisson error structure}
\label{fig:predict}
\end{figure}

\begin{figure}
\begin{center}
     \includegraphics[width=0.9\textwidth]{Cases_regmodel_resid.pdf}
\end{center}
\caption{Histogram of residuals from the regression model (\autoref{eq:reg}) and a quasipoisson error structure}
\label{fig:resid}
\end{figure}

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Sun Mar 22 17:09:42 2015
\begin{tabular}{lrrrrrr}
  \hline
 & Df & Deviance & Resid. Df & Resid. Dev & F & Pr($>$F) \\ 
  \hline
NULL &  &  & 59 & 1532.80 &  &  \\ 
  Age & 4 & 1345.17 & 55 & 187.63 & 204.01 & 0.0000 \\ 
  Ethnicity & 3 & 19.88 & 52 & 167.75 & 4.02 & 0.0141 \\ 
  NZDep & 2 & 17.10 & 50 & 150.65 & 5.19 & 0.0102 \\ 
  Age:Ethnicity & 12 & 80.80 & 38 & 69.85 & 4.09 & 0.0004 \\ 
   \hline
\end{tabular}\caption{Significance of different predictor variables for measles risk factors from 2007-2014, estimated using \autoref{eq:reg} and a quasipoisson error structure}
\label{table:anova}
\end{table}

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Sun Mar 22 17:09:42 2015
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -6.1644 & 0.1267 & -48.65 & 0.0000 \\ 
  Age2-4 & -1.0351 & 0.1972 & -5.25 & 0.0000 \\ 
  Age5-17 & -0.9954 & 0.1368 & -7.28 & 0.0000 \\ 
  Age18-24 & -1.7204 & 0.1926 & -8.93 & 0.0000 \\ 
  Age25+ & -3.1697 & 0.1622 & -19.55 & 0.0000 \\ 
  EthnicityAsian & 0.2464 & 0.3168 & 0.78 & 0.4415 \\ 
  EthnicityMaori & 0.2081 & 0.2121 & 0.98 & 0.3327 \\ 
  EthnicityPacific & 1.2200 & 0.2134 & 5.72 & 0.0000 \\ 
  NZDep4-7 & -0.2609 & 0.0948 & -2.75 & 0.0090 \\ 
  NZDep8-10 & -0.3043 & 0.1040 & -2.93 & 0.0058 \\ 
  Age2-4:EthnicityAsian & -2.3148 & 1.3319 & -1.74 & 0.0903 \\ 
  Age5-17:EthnicityAsian & -1.2453 & 0.4566 & -2.73 & 0.0096 \\ 
  Age18-24:EthnicityAsian & -1.0187 & 0.5743 & -1.77 & 0.0841 \\ 
  Age25+:EthnicityAsian & 0.2344 & 0.4298 & 0.55 & 0.5887 \\ 
  Age2-4:EthnicityMaori & -0.1013 & 0.3644 & -0.28 & 0.7826 \\ 
  Age5-17:EthnicityMaori & -0.1032 & 0.2521 & -0.41 & 0.6846 \\ 
  Age18-24:EthnicityMaori & -0.5723 & 0.4409 & -1.30 & 0.2022 \\ 
  Age25+:EthnicityMaori & -1.0349 & 0.5127 & -2.02 & 0.0506 \\ 
  Age2-4:EthnicityPacific & -0.9798 & 0.4676 & -2.10 & 0.0429 \\ 
  Age5-17:EthnicityPacific & -1.5709 & 0.3312 & -4.74 & 0.0000 \\ 
  Age18-24:EthnicityPacific & -1.0354 & 0.5003 & -2.07 & 0.0453 \\ 
  Age25+:EthnicityPacific & -0.6629 & 0.4279 & -1.55 & 0.1296 \\ 
   \hline
\end{tabular}\caption{Summary of the regression model results, estimated using \autoref{eq:reg} and a quasipoisson error structure, for measles risk factors for cases from 2007-2014}
\label{table:regression_riskfactors}
\end{table}
\vspace{5mm} %5mm vertical space

Apart from over-representation of some MLA categories discussed above and not included here, the results of the regression model suggest that age is a strong predictor of being a measles case (\autoref{table:regression_riskfactors}). Indeed, all age categories are significantly less likely to be measles cases compared to 0--1 year olds, and the likelihood generally decreases with age (\autoref{table:regression_riskfactors}, \autoref{fig:cases_all}, \autoref{fig:percap_all}, see also \autoref{fig:ageinyears}).

The majority of measles cases are among people of European ancestry (\autoref{table:percap} and \autoref{fig:cases_all}). Indeed, the majority of cases were in more wealthy European 5--17 year olds (\autoref{fig:cases_all}). However, per capita rates for this large age class were lower than in the very young (0--1 year olds, \autoref{fig:percap_all}). People of Pacific origin are over-represented as measles cases on a per capita basis ($\beta$ = 1.22, standard error (SE) = 0.21, p--value < 0.0001; \autoref{table:regression_riskfactors}), and in particular the 0--1 year old less wealthy (NZDep 8--10) (\autoref{fig:percap_all}. Note however that generally those people in NZDep levels 4--10 are less represented on a per capita basis (NZDep 4--7 $\beta$ = -0.26, SE = 0.09, p--value 0.009; NZDep 8--10 $\beta$ = -0.3, SE = 0.1, p--value == 0.006; \autoref{table:regression_riskfactors}), and in European and Maori ethnicities there is a trend in the opposite way than Pacific peoples, with the more wealthy more likely to be cases (\autoref{fig:percap_all} and \autoref{fig:cases_all}).

There are some other age:ethnicity classes that are significantly less represented in the data compared to Europeans in those ages classes, particularly in the 5--17 age classes. In later measles outbreaks (since 2007) there has been a shift in the distribution of ages infected. The very young (<2 years old) are still most likely to be infected, but school aged children and older teenagers are more likely to be represented then the under tens (\autoref{fig:ageinyears} and \autoref{fig:ageandvac}). This pattern suggests that improving vaccination coverage in the young is reducing the burden of measles in those age categories (\autoref{table:percap},\autoref{table:sero}, and \autoref{sub:popim}). Interestingly the regression results suggest risk of measles in the 5--17 year age category was greater for Europeans and Maori (\autoref{table:percap}, \autoref{table:regression_riskfactors}, \autoref{fig:percap_all}).

@@

\begin{table}
%latex.default(incidence, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{incidence}\tabularnewline
\hline
Georgia&$1806$\tabularnewline
Namibia&$ 455$\tabularnewline
Equatorial Guinea&$ 436$\tabularnewline
Angola&$ 409$\tabularnewline
Lebanon&$ 379$\tabularnewline
Nigeria&$ 313$\tabularnewline
Somalia&$ 311$\tabularnewline
Sudan&$ 260$\tabularnewline
Lesotho&$ 251$\tabularnewline
Uganda&$ 217$\tabularnewline
Netherlands&$ 159$\tabularnewline
Central African Republic&$ 132$\tabularnewline
Turkey&$ 100$\tabularnewline
Sri Lanka&$ 100$\tabularnewline
Togo&$  85$\tabularnewline
Sudan&$  76$\tabularnewline
Gabon&$  75$\tabularnewline
Niger&$  71$\tabularnewline
Nepal&$  68$\tabularnewline
Benin&$  63$\tabularnewline
Ethiopia&$  57$\tabularnewline
Romania&$  51$\tabularnewline
Pakistan&$  49$\tabularnewline
Thailand&$  40$\tabularnewline
Indonesia&$  39$\tabularnewline
Qatar&$  36$\tabularnewline
Cameroon&$  35$\tabularnewline
Syrian Arab Republic&$  34$\tabularnewline
Djibouti&$  33$\tabularnewline
United Kingdom&$  31$\tabularnewline
\hline
\end{tabular}\end{center}\caption{Highest measles incidence per million (2013)}
\label{table:incidence12}
\end{table}

\begin{table}
%latex.default(vaccinecover, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{cover}\tabularnewline
\hline
Central African Republic&$29$\tabularnewline
Somalia&$29$\tabularnewline
Equatorial Guinea&$41$\tabularnewline
South Sudan&$55$\tabularnewline
Lesotho&$61$\tabularnewline
Pakistan&$61$\tabularnewline
Vanuatu&$65$\tabularnewline
Gabon&$70$\tabularnewline
Marshall Islands&$70$\tabularnewline
Papua New Guinea&$70$\tabularnewline
Timor-Leste&$70$\tabularnewline
Mali&$72$\tabularnewline
Togo&$72$\tabularnewline
Nigeria&$73$\tabularnewline
Liberia&$74$\tabularnewline
San Marino&$74$\tabularnewline
Eritrea&$75$\tabularnewline
Paraguay&$75$\tabularnewline
Solomon Islands&$76$\tabularnewline
Iraq&$77$\tabularnewline
Ethiopia&$78$\tabularnewline
Haiti&$78$\tabularnewline
Yemen&$78$\tabularnewline
Chad&$79$\tabularnewline
Kenya&$79$\tabularnewline
Congo&$80$\tabularnewline
Djibouti&$80$\tabularnewline
Mauritania&$80$\tabularnewline
Zambia&$80$\tabularnewline
Afghanistan&$82$\tabularnewline
\hline
\end{tabular}\end{center}\caption{Lowest national measles vaccine cover (\%), 2013}
\label{table:cover12}
\end{table}

\begin{table}
%latex.default(im, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{immigration}\tabularnewline
\hline
Australia&$130703$\tabularnewline
China&$ 26380$\tabularnewline
United Kingdom&$ 25155$\tabularnewline
Japan&$ 20534$\tabularnewline
United States&$ 20247$\tabularnewline
Germany&$ 14793$\tabularnewline
France&$ 12173$\tabularnewline
Korea, Republic of&$ 10398$\tabularnewline
India&$  7013$\tabularnewline
Fiji&$  6040$\tabularnewline
Malaysia&$  5981$\tabularnewline
Thailand&$  5233$\tabularnewline
Canada&$  5013$\tabularnewline
Singapore&$  4887$\tabularnewline
Samoa&$  4047$\tabularnewline
South Africa&$  3763$\tabularnewline
Tonga&$  3477$\tabularnewline
Hong Kong&$  3281$\tabularnewline
Philippines&$  3147$\tabularnewline
Taiwan&$  2915$\tabularnewline
Indonesia&$  2580$\tabularnewline
Netherlands&$  2509$\tabularnewline
Brazil&$  2265$\tabularnewline
Sweden&$  1862$\tabularnewline
Denmark&$  1605$\tabularnewline
Switzerland&$  1376$\tabularnewline
Ireland&$  1331$\tabularnewline
Chile&$  1157$\tabularnewline
Saudi Arabia&$  1141$\tabularnewline
Russia&$  1068$\tabularnewline
\hline
\end{tabular}\end{center}\caption{Non-New Zealander travel and immigration numbers, 0--19 year olds, 2013}
\label{table:imtravel12}
\end{table}

\begin{table}
%latex.default(nz, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{immigration}\tabularnewline
\hline
Australia&$1017540$\tabularnewline
United States&$ 142200$\tabularnewline
Fiji&$ 112260$\tabularnewline
United Kingdom&$  96680$\tabularnewline
China&$  73340$\tabularnewline
Cook Islands&$  72820$\tabularnewline
Samoa&$  44060$\tabularnewline
Thailand&$  42060$\tabularnewline
India&$  40760$\tabularnewline
Indonesia&$  25100$\tabularnewline
Japan&$  21840$\tabularnewline
Canada&$  21720$\tabularnewline
Tonga&$  20120$\tabularnewline
Singapore&$  18300$\tabularnewline
Hong Kong&$  18000$\tabularnewline
Malaysia&$  17480$\tabularnewline
Philippines&$  17260$\tabularnewline
France&$  15980$\tabularnewline
Vanuatu&$  15580$\tabularnewline
South Africa&$  14000$\tabularnewline
Korea, Republic of&$  13980$\tabularnewline
Italy&$  13620$\tabularnewline
Germany&$  12800$\tabularnewline
Viet Nam&$  12640$\tabularnewline
Taiwan&$  10300$\tabularnewline
New Caledonia&$   7620$\tabularnewline
Spain&$   6560$\tabularnewline
Netherlands&$   6260$\tabularnewline
Papua New Guinea&$   6200$\tabularnewline
Ireland&$   6160$\tabularnewline
\hline
\end{tabular}\end{center}\caption{New Zealander travel numbers by destination, 2013}
\label{table:nztravel12}
\end{table}
